\vspace{-2cm}A generative model for urban firm networks

### 2022-05-23 CoverLetter.tex
submission_ok

\recipient{Editors PLoS ONE}{}
\date{May 12th 2022}
\opening{Dear Editors,}
\closing{Yours faithfully,\\
Juste Raimbault\\
LASTIG, Univ Gustave Eiffel, IGN-ENSG
}

@@

\justify
We are pleased, with my co-authors Natalia Zdanowska (Luxembourg Institute of Socio-economic Research) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modeling growth of urban firm networks'' for consideration for publication in PLoS ONE. This article introduces a generative model for urban firm networks, to understand underlying processes, with possible applications to public policies.

>>>>

\recipient{Editors PLoS ONE}{}
\date{May 23th 2022}
\opening{Dear Editors,}
\closing{Yours faithfully,\\
Juste Raimbault\\
LASTIG, Univ Gustave Eiffel, IGN-ENSG
}

@@

\justify
We are pleased, with my co-authors Natalia Zdanowska (Luxembourg Institute of Socio-economic Research) and Elsa Arcaute (University College London), to submit an original research article entitled ``Modelling growth of urban firm networks'' for consideration for publication in PLoS ONE. This article introduces a generative model for urban firm networks, to understand underlying processes, with possible applications to public policies.

### 2022-05-23 main.tex
submission_ok

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterised by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. In addition, strengthening city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}. This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also allow us to infer the consequences of unforeseen exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

@@

Secondly, we consider the geographical process of metropolisation. The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{sassen1991global}. Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most skilled workers and innovative industries \cite{RozenblatPumain2007}. Metropolisation gradients have been linked in the literature on systems of cities to several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}. We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities, or metropolises of regional importance, attracting the biggest amount of links and involved in a high number of interactions with other cities.  The later is computed using a Pearson correlation estimator on logarithms as follows:
\begin{equation}
    M = \hat{\rho} \left[\log (d_i) , \log (E_i) \right]
\end{equation}

@@

We run a specific experiment to study the impact of the urban hierarchy on firms dynamics. Rank-size law in cities is a central stylised fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}. I this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should help disentangling the variability due to model parameters from the variability due to the specific configuration of the urban system \cite{raimbault2019space}. We change the initial hierarchical exponent $\alpha$ from 0.1 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$. The behaviour of the internationalisation and metropolisation indicators are shown in Fig.~\ref{fig:fig7}. As expected, the hierarchy of the urban system has an important impact on model outcomes. We find that more hierarchical urban systems tend to produce more international networks (recall that the internationalisation indicator corresponds to the modularity, the latter is minimised when the networks span more across different countries). This is consistent with the existence of globalised metropolis \cite{sassen1991global}. This effect is mitigated by reducing the interaction ranges. Regarding metropolisation, we observe the existence of maxima as a function of the initial hierarchy across different interaction ranges $d_0$. This means that a too high hierarchy is detrimental to the global metropolisation when considering all cities, what may be due to the fact that in such a case very large cities are capturing all flows, leaving nothing to intermediate and medium-sized cities. Interestingly, the optimal value $\alpha \simeq 1.1$ is not far from the observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), implying that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.

>>>>

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterised by highly interconnected urban centres at the global level \cite{Sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. In addition, strengthening city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}. This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also allow us to infer the consequences of unforeseen exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

@@

Secondly, we consider the geographical process of metropolisation. The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{Sassen1991}. Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most skilled workers and innovative industries \cite{RozenblatPumain2007}. Metropolisation gradients have been linked in the literature on systems of cities to several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}. We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities, or metropolises of regional importance, attracting the biggest amount of links and involved in a high number of interactions with other cities.  The later is computed using a Pearson correlation estimator on logarithms as follows:
\begin{equation}
    M = \hat{\rho} \left[\log (d_i) , \log (E_i) \right]
\end{equation}

@@

We run a specific experiment to study the impact of the urban hierarchy on firms dynamics. Rank-size law in cities is a central stylised fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}. I this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should help disentangling the variability due to model parameters from the variability due to the specific configuration of the urban system \cite{raimbault2019space}. We change the initial hierarchical exponent $\alpha$ from 0.1 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$. The behaviour of the internationalisation and metropolisation indicators are shown in Fig.~\ref{fig:fig7}. As expected, the hierarchy of the urban system has an important impact on model outcomes. We find that more hierarchical urban systems tend to produce more international networks (recall that the internationalisation indicator corresponds to the modularity, the latter is minimised when the networks span more across different countries). This is consistent with the existence of globalised metropolis \cite{Sassen1991}. This effect is mitigated by reducing the interaction ranges. Regarding metropolisation, we observe the existence of maxima as a function of the initial hierarchy across different interaction ranges $d_0$. This means that a too high hierarchy is detrimental to the global metropolisation when considering all cities, what may be due to the fact that in such a case very large cities are capturing all flows, leaving nothing to intermediate and medium-sized cities. Interestingly, the optimal value $\alpha \simeq 1.1$ is not far from the observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), implying that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.
